Harmony: A Verifiable Systems Programming Language

### 2022-01-16 paper/paper.tex
Latest

@@

Harmony transforms the Kripke structure into a DAG of its strongly
connected components.  The states in the sink components of this
DAG can be considered the terminal states of the Harmony program.
States in which the context bag is empty represent states in which
all threads have terminated.  In the case the context bag is not
empty, we check if there are any non-\emph{eternal} contexts.
(Harmony programmers can
specify that certain threads are eternal, such as server threads
that operate in an infinite loop waiting for requests.) If
so, Harmony reports a \emph{deadlock}. If there
is no deadlock and a behavior automaton was specified, Harmony also
checks if the automaton is in a terminal state.  If not, a
\emph{behavior violation} is reported.

@@

A thread is considered \emph{blocked} if it can only make progress
(modify the state) if another thread makes a transition first.
This is not usually a problem: a thread waiting on a lock in a spinloop
is a common way of synchronizing threads, and Harmony does not report
such issues.  We call this \emph{passive busy waiting} as the thread
does not modify the state.

>>>>

In order to make deadlock detection possible, Harmony programs are
required to eventually terminate under fair selection of its
nondetermistic choices.  In other words, from each state there
should always exist an extension of the execution in which the
Harmony program terminates.  Harmony supports \emph{eternal}
threads---threads that never terminate---for example to model
servers.  A Harmony program is considered terminated if it is in a
state in which all its remaining threads are eternal and \emph{blocked}.
A thread is considered blocked if it can only update the shared
state after another thread has updated the shared state (for example,
a thread waiting on a spinlock).

@@

Harmony transforms the Kripke structure into a DAG of its strongly
connected components.  The states in the sink components of this
DAG can be considered the terminal states of the Harmony program.
States in which the context bag is empty represent states in which
all threads have terminated.  In the case the context bag is not
empty, we check if there are any non-\emph{eternal} contexts or
contexts that are not blocked.
If so, Harmony reports a \emph{deadlock}. If there
is no deadlock and a behavior automaton was specified, Harmony also
checks if the automaton is in a terminal state.  If not, a
\emph{behavior violation} is reported.

@@

As mentioned above, a thread is considered blocked if it can only
modify the shared state if another thread updates the shared state
first.  This is not usually a problem: a thread waiting on a lock
in a spinloop is a common way of synchronizing threads, and Harmony
does not report such issues.  We call this \emph{passive busy
waiting} as the thread does not modify the state.
